---
title: "Import microbial plate growth data"
author: "Maximilian Berthold, Douglas A. Campbell"
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
    code_folding: hide
---

This .Rmd imports Molecular Device Fluorescence well plate files, generated using a single wavelength protocol, with one or more plates in the file.
The size of the well plate does not matter; the import creates columns for RowColumn LetterNumber (A01 etc.) for the largest plate type in the list of files.
Import of smaller plate types results in NA in the non-existent RowColumn for that plate.

```{r load libraries, echo=FALSE, message = FALSE, warning = FALSE} 
# libraries; Note check actual dependencies
library(tidyverse) #core tidyverse packages
library(lubridate) #tidyverse dates
library(googledrive) #accessing googledrive
library(googlesheets4) #accessing googlesheets
library(minpack.lm) #curve fitting
library(broom) #tidy model outputs

```

```{r variable names for file import & processing MOLECULAR DEVICES}
Project <- "BIOL2201"

#deauthorizes access to googlesheet
gs4_deauth()

#read contents of MetaData from GoogleSheet
MetaData <- read_sheet("https://docs.google.com/spreadsheets/d/1FTKwYlJXd0ze9WwVXCvtwe5qkpN3_HNbeeM3hg7xnrI/edit#gid=0")

#set variables for file import & processing
DataPathMD <- file.path("RawData", "Single600Files", fsep = .Platform$file.sep)
file_id <- ".txt"

DataOut <- "ProcessData"

FileEncodeMD <- "UTF-16LE" 
DelimiterMD <- "\t"
HeaderRowsMD <- 2

#list of files
OD_files <- list.files(path = DataPathMD, pattern = file_id, full.names = TRUE)

OD_files
#test for duplicate file names in chl_files
unique(duplicated(OD_files))

```

```{r guess encoding}
guess_encoding(file = OD_files[1])

```

```{r read files, warning = FALSE, message = FALSE}
#"id" parameter of read_delim might replace read_delim_plus
#a read function using tidyverse::read_delim that skips a fixed number of header rows, and adds columns to the dataframe containing the filename and the file creation date time.
# 
read_delim_plus <- function(flnm, delimiter, headerrows, fileencode){read_delim(flnm, delim = delimiter,  col_names = TRUE,  skip = headerrows, escape_double = FALSE,  locale = locale(encoding = fileencode), trim_ws = TRUE) %>%
    mutate(Filename = flnm)
  }


OD_data <- OD_files %>%
  map_df(~read_delim_plus(flnm = ., delimiter = DelimiterMD,  headerrows = HeaderRowsMD,  fileencode = FileEncodeMD))

```

```{r tidy OD_data}
OD <- OD_data %>%
  select(-c("...1", "Temperature(¡C)", "...27")) %>%
  filter(!is.na(A1))  %>%
  separate(Filename, into = c("Path1", "Path2", "DDMMYYYY", "HHMM","Group", "Bench", "Temp_C", "txt"), sep = "([\\/\\_\\.])", remove = FALSE) %>%
  select(-c("Path1", "Path2", "txt")) %>%
  relocate(c("DDMMYYYY", "HHMM","Group", "Bench", "Temp_C")) %>%
  unite(DDMMYYYY_HHMM,  DDMMYYYY:HHMM, sep = "_", remove = TRUE) %>%
  mutate(DDMMYYYY_HHMM = dmy_hm(DDMMYYYY_HHMM))

#3 rows of data to average from each well measurement

OD <- OD %>%
  group_by(DDMMYYYY_HHMM, Group, Bench, Temp_C, Filename) %>%
  summarize(across(A1:D6, mean)) %>%
  ungroup()
  
```


```{r E_hours}
OD <- OD %>%
  group_by(Group, Bench, Temp_C) %>%
  mutate(E_hours =  as.numeric((DDMMYYYY_HHMM - min(DDMMYYYY_HHMM))) / 3600, .after = DDMMYYYY_HHMM) %>%
  ungroup()
```

```{r OD test plot}
OD %>% ggplot() +
  geom_point(aes(x = E_hours, y = A1)) + 
  facet_grid(rows = vars(Group), cols = vars(Bench, Temp_C)) +
  theme_bw()

```

Merge OD with MetaData

```{r long format}
OD_long <- OD %>%
  pivot_longer(cols = -c(DDMMYYYY_HHMM, E_hours, Group, Bench, Temp_C, Filename), names_to = "Well", values_to = "OD_600") %>%
  separate(Well, into = c("Row", "Col"), sep = 1, remove = FALSE)

```


```{r logOD}
OD_long <- OD_long %>%
  mutate(logOD_600 = log(OD_600))
```

```{r OD facet plot}
OD_long %>% 
  filter(Temp_C == 37,
         Group == "TUES") %>%
  ggplot() +
  geom_point((aes(x = E_hours, y = logOD_600, colour = Row))) + 
  facet_grid(rows = vars(Bench), cols = vars(Col)) +
  theme_bw()
```

```{r save GrowthLong, echo=FALSE}
saveRDS(OD_long, file.path(DataOut, 
paste(Project, "OD_long.Rds", sep = "_"), fsep = .Platform$file.sep))
```


Implement Growth Curve fits well by well.
Define equations as functions.
x will be taken from 'E_days' when we run the fit.
```{r logistic_eqn}
LogisticEqn <-  function(x, Pmax, Mu, Intercept){(Pmax*Intercept*exp(Mu*x))/(Pmax + (Intercept*(exp(Mu*x)-1)))
}

possibnlSLM = possibly(.f = nlsLM, otherwise = NULL)
```

Fit logistic growth trajectories using nest purrr:map & broom::augment using start, lower & upper settings extracted from each nest on the fly.
Extracting nest-specific start, lower & upper settings from each nest on the fly may be necessary if the 'nests' contain diverse data patterns that fail to fit with generic start, lower & upper parameters extracted from the entire dataset
```{r growth fits}
OD_nest <- OD_long %>%
  filter(Group == "TUES") %>%
  nest(data = c(DDMMYYYY_HHMM, Filename, E_hours, OD_600, logOD_600))

OD_nest <- OD_nest %>%
  mutate(ExpInit_h = as.numeric(map(data, ~((.$logOD_600[2]) - (.$logOD_600[1]))/.$E_hours[2]))) %>%
  mutate(FitLog = map(data, ~possibnlSLM(OD_600 ~ LogisticEqn(x = E_hours, Intercept, Mu, Pmax),
                                         data = .x,
                                         start = list(Intercept = min(.$OD_600, na.rm = TRUE), Mu = 0.1, Pmax = max(.$OD_600, na.rm = TRUE))
                                         )
  ),
  TidiedLog = map(FitLog, tidy),
  PredictLog = map(FitLog, augment)
  )

  # ParamLog = map(FitLog, glance)
  # )

OD_nest %>%
  filter(Bench == "B1",
         Col %in% c(2,3,5)) %>%
  unnest(PredictLog) %>%
  ggplot() + 
  geom_point(aes(x = E_hours, y = OD_600)) + 
  geom_ribbon(aes(x = E_hours, ymin =  `.fitted` - `.resid`, ymax =  `.fitted` + `.resid`), colour = "grey", fill = "grey") +
  facet_grid(col = vars(Col), rows = vars(Temp_C)) +
  theme_bw()

Mu_est <- OD_nest %>% 
  dplyr::select(-c(data, FitLog, PredictLog)) %>%
  unnest(TidiedLog) %>%
  select(-statistic) %>%
  pivot_wider(names_from = term, values_from = c(estimate:p.value))

Mu_est %>%
  # filter(Bench == "B1",
  #        Col %in% c(2,3,5)) %>%
  ggplot() + 
  geom_point(aes(x = estimate_Mu, y = ExpInit_h, size = estimate_Pmax)) +
  #coord_fixed(ratio = 1, xlim = c(0, 0.2), ylim = c(0, 0.2)) +
  facet_grid(cols = vars(Bench), rows = vars(Temp_C)) +
  theme_bw()
  
```


Implement Growth Curve fits pool by pool of 4 rows together from one column.

Fit treatment specific logistic growth trajectories using nest purrr:map & broom::augment.
Using nest specific start, lower & upper settings extracted from each nest on the fly. may be necessary if the 'nests' contain diverse data patterns that fail to fit with generic start, lower & upper parameters extracted from the entire dataset
```{r growth fits}
OD_pool_nest <- OD_long %>%
  filter(Group == "TUES") %>%
  nest(data = c(DDMMYYYY_HHMM, Well, Row, Filename, E_hours, OD_600, logOD_600))

OD_pool_nest <- OD_pool_nest %>%
  mutate(FitLog = map(data, ~possibnlSLM(OD_600 ~ LogisticEqn(x = E_hours, Intercept, Mu, Pmax),
                                         data = .x,
                                         start = list(Intercept = min(.$OD_600, na.rm = TRUE), Mu = 0.1, Pmax = max(.$OD_600, na.rm = TRUE))
                                         )
  ),
  TidiedLog = map(FitLog, tidy),
  PredictLog = map(FitLog, augment)
  )

OD_pool_nest %>%
  filter(Bench == "B1",
         Col %in% c(2,3,5)) %>%
  unnest(PredictLog) %>%
  ggplot() + 
  geom_point(aes(x = E_hours, y = OD_600)) + 
  geom_ribbon(aes(x = E_hours, ymin =  `.fitted` - `.resid`, ymax =  `.fitted` + `.resid`), colour = "grey", fill = "grey") +
  facet_grid(col = vars(Col), rows = vars(Temp_C)) +
  theme_bw()

  
```