---
title: "Fit Microbial Growth Curves Using Microbial WellPlate Data"
author: "Maximilian Berthold, Douglas A. Campbell"
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
    code_folding: hide
editor_options: 
  markdown: 
    wrap: sentence
---

Enter Your Name
```{r enter name}
LastName <- "Campbell"
FirstName <- "Douglas"

LastName_FirstName <- paste(LastName, FirstName, sep = "_")
```

Enter your LabDay
```{r enter LabDay}
#TUES, WED, THURS, FRI
LabDay <- "TUES"
```

# Report by `r LastName_FirstName`

# Introduction
This .Rmd uses an imported .Rds containing tidied data from Molecular Device Fluorescence well plate files, generated using multiple excitation/emission wavelength pairs, along with MetaData imported from an .xlsx file.

It fits logistic growth curves to data grouped by SAMPLE (taxa) and LIGHT.


# Materials & Methods

```{r load libraries, echo=FALSE, message = FALSE, warning = FALSE}
# libraries; Note check actual dependencies
library(tidyverse) #core tidyverse packages
library(minpack.lm) #curve fitting
library(broom) #tidy model outputs
library(photobiology) #photobiology::w_length2rgb
library(kableExtra)

```


```{r variable names for processing}

Project <- "BIOL2201"

#File Ex Em nm settings lost from files during import b/c of Header structure, added back during tidying.

Em <- 680
Ex <- c(440, 520, 620)

#Add colour coding vector
ExNM = (photobiology::w_length2rgb(Ex))
names(ExNM) <- Ex

```


```{r set variables for file import, export}

file_id = ".Rds"  #set value to search for file type

#Read in .Rds, stored on OneDrive; file path is fussy and contains spaces & " - Mount Allison University"
#Remember to force sync to local harddrive b/f attempting import into RStudio

#Currently read in and process each LabDay separately; Alternative could use recursive read in to accumulate files from all LabDays together, then filter upon output?

#DataInOneDrive <- paste("~/OneDrive - Mount Allison University/BIOL2201_2024/WI2024Labs/LAB3_MicroGrowth/StudentData", sep = "")

#list of files from OneDrive
#ExEm_meta_files <- list.files(path = DataInOneDrive, pattern = file_id, full.names = TRUE)

#Import from local OneDrive path does not work from Posit.cloud.  Need to move data files to git tracked project, push to GitHub, pull into Posit.cloud
#List files stored on GitHub
ExEm_meta_files <- list.files(path = file.path("StudentData"), pattern = file_id, full.names = TRUE)

ExEm_meta_files

```

## Read in Fluorescence data and MetaData tidied and combined in another script
```{r read ExEm_meta file, error = FALSE, message = FALSE, warning = FALSE}

ExEm_meta <-readRDS(file = paste("StudentData/BIOL2201_", LabDay, "_ExEm_meta.Rds", sep = ""))
```

## Check Data with a Test Plot

```{r test plot}
ExEm_meta |>
  filter(Ex_nm == 440) |>
  ggplot() +
  geom_point(aes(x = E_hours, y = Fluor, colour = INITIAL)) +
  #scale_colour_manual(values = ExNM) +
  facet_grid(rows = vars(SAMPLE), cols = vars(LIGHT)) +
  labs(title = LabDay, subtitle = paste(Ex[1], "_nm"))+
  theme_bw()

ExEm_meta |>
  ggplot() +
  geom_point(aes(x = E_hours, y = Fluor, colour = LIGHT)) +
  #scale_colour_manual(values = ExNM) +
  facet_grid(rows = vars(SAMPLE), cols = vars(Ex_nm)) +
  labs(title = LabDay) +
  theme_bw()
```

# Results
Implement MultiSpectral Analyses and Growth Curve fits well by well.

Define fitting logistic equation as function for R.

```{r logistic_eqn}
#x will be taken from 'E_hours' when we run the fit.
LogisticEqn <-  function(x, Pmax, Mu, Intercept){(Pmax*Intercept*exp(Mu*x))/(Pmax + (Intercept*(exp(Mu*x)-1)))
}

#ModGompertzEqn <- function(x, Gmax,Gmu,Lag){(Gmax*(exp(-exp((Gmu*exp(1))/Gmax*(Lag-x)+1))))}

possibnlSLM = possibly(.f = nlsLM, otherwise = NULL)
```

Fit Sample & Treatment specific logistic growth trajectories using nest purrr:map & broom::augment using nest specific start, lower & upper settings extracted from each data nest on the fly.
This may be necessary if the 'nests' contain diverse data patterns that fail to fit with generic start, lower & upper parameters extracted from the entire dataset

```{r growth fits}
ExEm_nest <- ExEm_meta |>
  filter(Fluor >= 0,
         Fluor <= 1000) |> #hack fix to remove bad points from 620 nm Ex, 20 h)
  nest(data = -c(DAY, LIGHT, Ex_nm, SAMPLE, PAR_uE, Temp_C))

ExEm_nest <- ExEm_nest|> 
  mutate(FitLog = map(data, ~possibnlSLM(Fluor ~ LogisticEqn(x = E_hours, Intercept, Mu, Pmax),
                                         data = .x,
                                         start = list(Intercept = min(.$Fluor, na.rm = TRUE), Mu = 0.1, Pmax = max(.$Fluor, na.rm = TRUE))
                                         )
  ),
  PredictLog = map(FitLog, augment),
  TidiedLog = map(FitLog, tidy),
  ParamLog = map(FitLog, glance)
  )
```

## Plotting Growth and Growth Parameters Across Strains (Samples) and Growth Light `r LabDay`

```{r plot growth parameters}
ExEm_nest |>
  unnest(PredictLog) |>
  ggplot() + 
  geom_point(aes(x = E_hours, y = Fluor, colour = as.factor(Ex_nm))) + 
  geom_line(aes(x = E_hours, y = `.fitted`, colour = as.factor(Ex_nm))) +
  geom_point(aes(x = E_hours, y =  `.resid`), size = 0.1) +
  scale_colour_manual(values = ExNM) +
  facet_grid(row = vars(SAMPLE), col = vars(LIGHT)) +
  labs(title = LabDay) +
  theme_bw()

ExEm_nest |> 
  dplyr::select(-c(data, FitLog, PredictLog, ParamLog)) |>
  unnest(TidiedLog) |>
  select(-statistic) |>
  pivot_wider(names_from = term, values_from = c(estimate:p.value)) |>
  ggplot() +
  geom_point(aes(x = Ex_nm, y = estimate_Mu, colour = LIGHT)) +
  facet_grid(rows = vars(SAMPLE)) +
  labs(title = LabDay) + 
  theme_bw()

ExEm_nest |> 
  dplyr::select(-c(data, FitLog, PredictLog, ParamLog)) |>
  unnest(TidiedLog) |>
  select(-statistic) |>
  pivot_wider(names_from = term, values_from = c(estimate:p.value)) |>
  ggplot() +
  geom_point(aes(x = Ex_nm, y = estimate_Mu, colour = LIGHT)) +
  geom_errorbar(aes(x = Ex_nm, ymin = estimate_Mu - std.error_Mu, ymax = estimate_Mu + std.error_Mu, colour = LIGHT)) +
  facet_grid(rows = vars(SAMPLE)) +
  labs(title = LabDay) + 
  theme_bw()

ExEm_nest |> 
  dplyr::select(-c(data, FitLog, PredictLog, ParamLog)) |>
  unnest(TidiedLog) |>
  select(-statistic) |>
  pivot_wider(names_from = term, values_from = c(estimate:p.value)) |>
  ggplot() +
  geom_point(aes(x = Ex_nm, y = estimate_Pmax, colour = LIGHT)) +
  facet_grid(rows = vars(SAMPLE)) +
  labs(title = LabDay) + 
  theme_bw()

ExEm_nest |> 
  dplyr::select(-c(data, FitLog, PredictLog, ParamLog)) |>
  unnest(TidiedLog) |>
  select(-statistic) |>
  pivot_wider(names_from = term, values_from = c(estimate:p.value)) |>
  ggplot() +
  geom_point(aes(x = Ex_nm, y = estimate_Pmax, colour = LIGHT)) +
  geom_errorbar(aes(x = Ex_nm, ymin = estimate_Pmax - std.error_Pmax, ymax = estimate_Pmax + std.error_Pmax, colour = LIGHT)) +
  facet_grid(rows = vars(SAMPLE)) +
  labs(title = LabDay) + 
  theme_bw()

```

## Table of Logistic Fit Parameters for `r LabDay`
```{r logistic fit parameters}

ExEm_nest |> 
  dplyr::select(-c(data, FitLog, PredictLog, ParamLog)) |>
  unnest(TidiedLog) |>
  select(-c(p.value, statistic)) |>
  pivot_wider(names_from = term, values_from = c(estimate:std.error)) |>
  rename(Intercept = estimate_Intercept,
         mu = estimate_Mu,
         Pmax = estimate_Pmax) |>
  mutate(across(where(is.numeric), \(x) round(x, 1)))

# kable(ExEm_nest |> 
#   dplyr::select(-c(data, FitLog, PredictLog, ParamLog)) |>
#   unnest(TidiedLog) |>
#   select(-statistic) |>
#   pivot_wider(names_from = term, values_from = c(estimate:p.value)) |>
#   rename(Intercept = estimate_Intercept,
#          mu = estimate_Mu,
#          Pmax = estimate_Pmax) |>
#   mutate(across(where(is.numeric), round, 1)),
#   digits = 1,
#   caption = paste(LabDay, "Logistic Growth Parameters", sep = " "))

```

# Discussion