---
title: "Fit Microbial Growth Curves Using Microbial WellPlate Data"
author: "Maximilian Berthold, Douglas A. Campbell"
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
    code_folding: hide
editor_options: 
  markdown: 
    wrap: sentence
---

Enter Your Name
```{r enter name}
LastName <- "Campbell"
FirstName <- "Douglas"

LastName_FirstName <- paste(LastName, FirstName, sep = "_")
```

Enter your LabDay
```{r enter LabDay}
#TUES, WED, THURS, FRI
LabDay <- "TUES"
```

# Report by `r LastName_FirstName`

# Introduction
This .Rmd uses an imported .Rds containing tidied data from Molecular Device Fluorescence well plate files, generated using multiple excitation/emission wavelength pairs, along with MetaData imported from an .xlsx file.

It fits logistic growth curves to data grouped by SAMPLE (taxa) and LIGHT.

# Materials & Methods

```{r load libraries, echo=FALSE, message = FALSE, warning = FALSE}
# libraries; Note check actual dependencies
library(tidyverse) #core tidyverse packages
library(minpack.lm) #curve fitting
library(broom) #tidy model outputs

```


```{r variable names for processing}

Project <- "BIOL2201"

#File Ex Em nm settings lost from files during import b/c of Header structure, added back during tidying.

Em <- 680
Ex <- c(440, 520, 620)

#Add colour coding vector
ExNM = (photobiology::w_length2rgb(Ex))
names(ExNM) <- Ex
```


```{r list files for import}

file_id = ".Rds"  #set value to search for file type

#Read in .Rds, stored on OneDrive; file path is fussy and contains spaces & " - Mount Allison University"
#Remember to force sync to local harddrive b/f attempting import into RStudio

#Currently read in and process each LabDay separately; Alternative could use recursive read in to accumulate files from all LabDays together, then filter upon output?

#DataInOneDrive <- paste("~/OneDrive - Mount Allison University/BIOL2201_2024/WI2024Labs/LAB3_MicroGrowth/StudentData", sep = "")

#list of files from OneDrive
#ExEm_meta_files <- list.files(path = DataInOneDrive, pattern = file_id, full.names = TRUE)

#Import from local OneDrive path does not work from Posit.cloud.  Need to move data files to git tracked project, push to GitHub, pull into Posit.cloud
#List files stored on GitHub
ExEm_meta_files <- list.files(path = file.path("StudentData"), pattern = file_id, full.names = TRUE)

ExEm_meta_files

```

## Read in Fluorescence data and MetaData, tidied and combined from data files, in another script
```{r read ExEm_meta file, error = FALSE, message = FALSE, warning = FALSE}

ExEm_meta <-readRDS(file = paste("StudentData/BIOL2201_", LabDay, "_ExEm_meta.Rds", sep = ""))
```

## Check Data with a Test Plot

```{r test plot}
ExEm_meta |>
  filter(Ex_nm == Ex[1]) |>
  ggplot() +
  geom_point(aes(x = E_hours, y = Fluor, colour = INITIAL)) +
  facet_grid(rows = vars(SAMPLE), cols = vars(LIGHT)) +
  labs(title = LabDay, subtitle = paste(Ex[1], "_nm"))+
  theme_bw()

ExEm_meta |>
  ggplot() +
  geom_point(aes(x = E_hours, y = Fluor, colour = LIGHT)) +
  facet_grid(rows = vars(SAMPLE), cols = vars(Ex_nm)) +
  labs(title = LabDay) +
  theme_bw()
```

# Results
Implement MultiSpectral Analyses and Growth Curve fits well by well.

Define fitting logistic equations as functions for R.

```{r logistic_eqns}
#x will be taken from 'E_hours' when we run the fit.
logistic_eqn <-  function(x, Pmax, Mu, Intercept){(Pmax*Intercept*exp(Mu*x))/(Pmax + (Intercept*(exp(Mu*x)-1)))
}

TestX = c(0:200)

plot(x = TestX,  y = logistic_eqn(x = TestX, Intercept = 10, Pmax = 100, Mu = 0.05))

#logistic_eqn with additional D 'death' parameter
logisticdeath_eqn <- function(x, Pmax, Mu, Intercept, D){((Pmax*Intercept*exp(Mu*x))/(Pmax + (Intercept*(exp(Mu*x)-1)))) - (x * D)
}
 
plot(x = TestX,  y = logisticdeath_eqn(x = TestX, Intercept = 10, Pmax = 100, Mu = 0.05, D = 0.05))

#ModGompertzEqn <- function(x, Gmax,Gmu,Lag){(Gmax*(exp(-exp((Gmu*exp(1))/Gmax*(Lag-x)+1))))}

possibnlSLM = possibly(.f = nlsLM, otherwise = NULL)
```

Fit Sample & Treatment specific logistic growth trajectories using nest purrr:map & broom::augment using nest specific start, lower & upper settings extracted from each data nest on the fly.
This may be necessary if the 'nests' contain diverse data patterns that fail to fit with generic start, lower & upper parameters extracted from the entire dataset

https://r4ds.hadley.nz/
```{r logistic growth fits}
ExEm_nest <- ExEm_meta |>
  filter(Fluor >= 0,
         Fluor <= 1000) |> #hack fix to remove bad points from 620 nm Ex, 20 h)
  nest(data = -c(DAY, LIGHT, Ex_nm, SAMPLE, PAR_uE, Temp_C))

ExEm_nest <- ExEm_nest|> 
  mutate(FitLog = map(data, ~possibnlSLM(Fluor ~ logistic_eqn(x = E_hours, Intercept, Mu, Pmax),
                                         data = .x,
                                         start = list(Intercept = min(.$Fluor, na.rm = TRUE), Mu = 0.1, Pmax = max(.$Fluor, na.rm = TRUE))
                                         )
  ),
  PredictLog = map(FitLog, augment),
  TidiedLog = map(FitLog, tidy),
  ParamLog = map(FitLog, glance)
  )
```

## Plotting Growth and Growth Parameters Across Strains (Samples) and Growth Light `r LabDay`

```{r plot data and logistic model fits}
ExEm_nest |>
  unnest(PredictLog) |>
  ggplot() + 
  geom_point(aes(x = E_hours, y = Fluor, colour = as.factor(Ex_nm))) + 
  geom_line(aes(x = E_hours, y = `.fitted`, colour = as.factor(Ex_nm))) +
  geom_point(aes(x = E_hours, y =  `.resid`), size = 0.1) +
  scale_colour_manual(values = ExNM) +
  facet_grid(row = vars(SAMPLE), col = vars(LIGHT)) +
  labs(title = LabDay) +
  theme_bw()

```


```{r plot logistic growth parameters}

ExEm_nest |> 
  dplyr::select(-c(data, FitLog, PredictLog, ParamLog)) |>
  unnest(TidiedLog) |>
  select(-statistic) |>
  pivot_wider(names_from = term, values_from = c(estimate:p.value)) |>
  ggplot() +
  geom_point(aes(x = Ex_nm, y = estimate_Mu, colour = LIGHT)) +
  facet_grid(rows = vars(SAMPLE)) +
  labs(title = LabDay) + 
  theme_bw()

ExEm_nest |> 
  dplyr::select(-c(data, FitLog, PredictLog, ParamLog)) |>
  unnest(TidiedLog) |>
  select(-statistic) |>
  pivot_wider(names_from = term, values_from = c(estimate:p.value)) |>
  ggplot() +
  geom_point(aes(x = Ex_nm, y = estimate_Mu, colour = LIGHT)) +
  geom_errorbar(aes(x = Ex_nm, ymin = estimate_Mu - std.error_Mu, ymax = estimate_Mu + std.error_Mu, colour = LIGHT)) +
  facet_grid(rows = vars(SAMPLE)) +
  labs(title = LabDay) + 
  theme_bw()

ExEm_nest |> 
  dplyr::select(-c(data, FitLog, PredictLog, ParamLog)) |>
  unnest(TidiedLog) |>
  select(-statistic) |>
  pivot_wider(names_from = term, values_from = c(estimate:p.value)) |>
  ggplot() +
  geom_point(aes(x = Ex_nm, y = estimate_Pmax, colour = LIGHT)) +
  facet_grid(rows = vars(SAMPLE)) +
  labs(title = LabDay) + 
  theme_bw()

ExEm_nest |> 
  dplyr::select(-c(data, FitLog, PredictLog, ParamLog)) |>
  unnest(TidiedLog) |>
  select(-statistic) |>
  pivot_wider(names_from = term, values_from = c(estimate:p.value)) |>
  ggplot() +
  geom_point(aes(x = Ex_nm, y = estimate_Pmax, colour = LIGHT)) +
  geom_errorbar(aes(x = Ex_nm, ymin = estimate_Pmax - std.error_Pmax, ymax = estimate_Pmax + std.error_Pmax, colour = LIGHT)) +
  facet_grid(rows = vars(SAMPLE)) +
  labs(title = LabDay) + 
  theme_bw()

```

## Table of Logistic Fit Parameters for `r LabDay`
```{r logistic fit parameters}

ExEm_nest |> 
  dplyr::select(-c(data, FitLog, PredictLog, ParamLog)) |>
  unnest(TidiedLog) |>
  select(-c(p.value, statistic)) |>
  pivot_wider(names_from = term, values_from = c(estimate:std.error)) |>
  rename(Intercept = estimate_Intercept,
         mu = estimate_Mu,
         Pmax = estimate_Pmax) |>
  mutate(across(where(is.numeric), \(x) round(x, 1)))


```

## Plot growth parameters that pass quality filter
```{r plot quality-filter growth parameters}

FilterFactor <- 2

ExEm_nest |> 
  dplyr::select(-c(data, FitLog, PredictLog, ParamLog)) |>
  unnest(TidiedLog) |>
  select(-statistic) |>
  pivot_wider(names_from = term, values_from = c(estimate:p.value)) |>
  filter(std.error_Mu < FilterFactor * estimate_Mu) |>
  ggplot() +
  geom_point(aes(x = Ex_nm, y = estimate_Mu, colour = LIGHT)) +
  geom_errorbar(aes(x = Ex_nm, ymin = estimate_Mu - std.error_Mu, ymax = estimate_Mu + std.error_Mu, colour = LIGHT)) +
  facet_grid(rows = vars(SAMPLE)) +
  labs(title = LabDay) + 
  theme_bw()


ExEm_nest |> 
  dplyr::select(-c(data, FitLog, PredictLog, ParamLog)) |>
  unnest(TidiedLog) |>
  select(-statistic) |>
  pivot_wider(names_from = term, values_from = c(estimate:p.value)) |>
  filter(std.error_Mu < FilterFactor * estimate_Mu) |>
  ggplot() +
  geom_point(aes(x = Ex_nm, y = estimate_Pmax, colour = LIGHT)) +
  geom_errorbar(aes(x = Ex_nm, ymin = estimate_Pmax - std.error_Pmax, ymax = estimate_Pmax + std.error_Pmax, colour = LIGHT)) +
  facet_grid(rows = vars(SAMPLE)) +
  labs(title = LabDay) + 
  theme_bw()

```

## Fit data with logisticdeath model to account for death phase?
```{r logistic death growth fits}
ExEmDeath_nest <- ExEm_meta |>
  filter(Fluor >= 0,
         Fluor <= 1000) |> #hack fix to remove bad points from 620 nm Ex, 20 h)
  nest(data = -c(DAY, LIGHT, Ex_nm, SAMPLE, PAR_uE, Temp_C))

ExEmDeath_nest <- ExEm_nest|> 
  mutate(FitLog = map(data, ~possibnlSLM(Fluor ~ logisticdeath_eqn(x = E_hours, Intercept, Mu, Pmax, D),
                                         data = .x,
                                         start = c(Intercept = min(.$Fluor, na.rm = TRUE), Mu = 0.1, Pmax = max(.$Fluor, na.rm = TRUE), D = 0.05),
                                         lower = c(Intercept = 0, Mu = 0, Pmax = 2, D = 0)
                                         )
  ),
  PredictLog = map(FitLog, augment),
  TidiedLog = map(FitLog, tidy),
  ParamLog = map(FitLog, glance)
  )
```


```{r plot data and logistic death model fits}
ExEmDeath_nest |>
  unnest(PredictLog) |>
  ggplot() + 
  geom_point(aes(x = E_hours, y = Fluor, colour = as.factor(Ex_nm))) + 
  geom_line(aes(x = E_hours, y = `.fitted`, colour = as.factor(Ex_nm))) +
  geom_point(aes(x = E_hours, y =  `.resid`), size = 0.1) +
  scale_colour_manual(values = ExNM) +
  facet_grid(row = vars(SAMPLE), col = vars(LIGHT)) +
  labs(title = LabDay) +
  theme_bw()

```

## Table of Logistic Death Fit Parameters for `r LabDay`
```{r logistic death fit parameters}

ExEmDeath_nest |> 
  dplyr::select(-c(data, FitLog, PredictLog, ParamLog)) |>
  unnest(TidiedLog) |>
  select(-c(p.value, statistic)) |>
  pivot_wider(names_from = term, values_from = c(estimate:std.error)) |>
  rename(Intercept = estimate_Intercept,
         mu = estimate_Mu,
         Pmax = estimate_Pmax,
         Death = estimate_D) |>
  mutate(across(where(is.numeric), \(x) round(x, 3)))

```

# Discussion

Discussion
1) Were growth rates across the different species different?  

2) Were growth rates different under HIGH vs. LOW light for each species?  

3) Were growth rates the same, or different, for the different pigments, within a species?  

	If the pigments show the same growth rates within a species, growth was 'balanced'; the level of each pigment increased at the same exponential (proportional) rate over time.  
	
	If the pigments show different growth rates  within a species, growth was 'imbalanced'; the level of the pigments is increasing at different rates, implying a change in cell composition over time.  
	

4) Does adding a Death parameter to the model better approximate your data?  

Report:  

Save this .Rmd file  

'Knit' to 'html' to generate a formatted output file.  
	Errors or code failures will halt the 'knit' output.
	The output should appear in the 'Files' tab, and as a popup; you may need to disable popup blockers.

Select the .html file from the 'Files' tab.  

Under the Gear icon select 'Export'  

Save the file.    

Rename the file to include your LastNameFirstName  

Submit the renamed .html to Moodle.

