---
title: "Import microbial plate growth data"
author: "Maximilian Berthold, Douglas A. Campbell"
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
    code_folding: hide
---

This .Rmd imports Molecular Device Fluorescence well plate files, generated using an absorbance scan protocol.
The size of the well plate does not matter; the import creates columns for RowColumn LetterNumber (A01 etc.) for the largest plate type in the list of files.
Import of smaller plate types results in NA in the non-existent RowColumn for that plate.

```{r load libraries, echo=FALSE, message = FALSE, warning = FALSE} 
# libraries; Note check actual dependencies
library(tidyverse) #core tidyverse packages
library(lubridate) #tidyverse dates
library(googledrive) #accessing googledrive
library(googlesheets4) #accessing googlesheets
library(minpack.lm) #curve fitting
library(broom) #tidy model outputs

```

```{r variable names for file import & processing MOLECULAR DEVICES}
Project <- "BIOL2201"

#deauthorizes access to googlesheet
gs4_deauth()

#read contents of MetaData from GoogleSheet
MetaData <- read_sheet("https://docs.google.com/spreadsheets/d/1FTKwYlJXd0ze9WwVXCvtwe5qkpN3_HNbeeM3hg7xnrI/edit#gid=0")

#set variables for file import & processing
DataPathMD <- file.path("RawData", "TestPlate", fsep = .Platform$file.sep)
file_id <- ".txt"

DataOut <- "ProcessData"

FileEncodeMD <- "UTF-16LE" 
DelimiterMD <- "\t"
HeaderRowsMD <- 2

#list of files
OD_files <- list.files(path = DataPathMD, pattern = file_id, full.names = TRUE)

OD_files
#test for duplicate file names in chl_files
unique(duplicated(OD_files))

```
```{r guess encoding}
guess_encoding(file = OD_files[1])

```

```{r read files}
#"id" parameter of read_delim might replace read_delim_plus
#a read function using tidyverse::read_delim that skips a fixed number of header rows, and adds columns to the dataframe containing the filename and the file creation date time.
# 
read_delim_plus <- function(flnm, delimiter, headerrows, fileencode){read_delim(flnm, delim = delimiter,  col_names = TRUE,  skip = headerrows, escape_double = FALSE,  locale = locale(encoding = fileencode), trim_ws = TRUE) %>%
    mutate(Filename = flnm)
  }


OD_data <- OD_files %>%
  map_df(~read_delim_plus(flnm = ., delimiter = DelimiterMD,  headerrows = HeaderRowsMD,  fileencode = FileEncodeMD))

```

```{r test spectra}
SpectraTest <- OD_data %>%
  filter(Wavelength >= 450,
         Wavelength <= 700) %>%
  filter(Filename == OD_files[6]) %>%
  ggplot() +
  geom_point(aes(x = Wavelength, y = A1)) + 
  facet_grid(cols = vars(Filename)) + 
  theme_bw()

SpectraTest

# ggsave(filename = file.path("Plots", "SpectraTest.png"), plot = SpectraTest)
```

```{r tidy OD_data}
OD <- OD_data %>%
  select(-c("Temperature(¡C)", "Basic Endpoint Protocol", "...27")) %>%
  filter(Wavelength == 600) %>%
  separate(Filename, into = c("Path1", "Path2", "YYYYMMDD", "HHMM","Plate", "Temp_C", "txt"), sep = "([\\/\\_\\.])", remove = FALSE) %>%
  select(-c("Path1", "Path2", "txt")) %>%
  relocate(c("YYYYMMDD", "HHMM","Plate", "Temp_C")) %>%
  unite(YYYYMMDD_HHMM,  YYYYMMDD:HHMM, sep = "_", remove = TRUE) %>%
  mutate(YYYYMMDD_HHMM = ymd_hm(YYYYMMDD_HHMM))

```

```{r E_hours}
OD <- OD %>%
  group_by(Plate, Wavelength) %>%
  mutate(E_hours =  as.numeric((YYYYMMDD_HHMM - min(YYYYMMDD_HHMM))) / 3600, .after = YYYYMMDD_HHMM) %>%
  ungroup()
```

```{r OD test plot}
OD %>% ggplot() +
  geom_line(aes(x = E_hours, y = A1)) + 
  theme_bw()

OD %>% ggplot() +
  geom_line(aes(x = E_hours, y = log(A1))) + 
  theme_bw()
```

Merge OD with MetaData
XXXX

```{r long format}
OD_long <- OD %>%
  pivot_longer(cols = -c(YYYYMMDD_HHMM:Wavelength, Filename), names_to = "Well", values_to = "OD_600") %>%
  separate(Well, into = c("Row", "Col"), sep = 1, remove = FALSE)

```


```{r logOD}
OD_long <- OD_long %>%
  mutate(logOD_600 = log(OD_600))
```

```{r OD facet plot}
OD_long %>% ggplot() +
  geom_line((aes(x = E_hours, y = OD_600))) + 
  facet_grid(rows = vars(Row), cols = vars(Col)) +
  theme_bw()

OD_long %>% ggplot() +
  geom_line((aes(x = E_hours, y = logOD_600))) + 
  facet_grid(rows = vars(Row), cols = vars(Col)) +
  theme_bw()
```


```{r save GrowthLong, echo=FALSE}
saveRDS(OD_long, file.path(DataOut, 
paste(Project, "OD_long.Rds", sep = "_"), fsep = .Platform$file.sep))
```

Implement Growth Curve fits well by well.
Define equations as functions.
x will be taken from 'E_days' when we run the fit.
```{r logistic_eqn}
LogisticEqn <-  function(x, Pmax, Mu, Intercept){(Pmax*Intercept*exp(Mu*x))/(Pmax + (Intercept*(exp(Mu*x)-1)))
}

ModGompertzEqn <- function(x, Gmax,Gmu,Lag){(Gmax*(exp(-exp((Gmu*exp(1))/Gmax*(Lag-x)+1))))}

possibnlSLM = possibly(.f = nlsLM, otherwise = NULL)
```

Fit treatment specific logistic growth trajectories using nest purrr:map & broom::augment using nest specific start, lower & upper settings extracted from each nest on the fly.
This may be necessary if the 'nests' contain diverse data patterns that fail to fit with generic start, lower & upper parameters extracted from the entire dataset
```{r growth fits}
#run nesting by passing in values of variables set at top; 
#!!sym(... allows nest to access the characters of the values passed in

OD_nest <- OD_long %>%
  nest(data = c(YYYYMMDD_HHMM, Filename, E_hours, OD_600, logOD_600))

#extract nest-specific start, lower & upper setting by passing in "."
OD_nest <- OD_nest %>% 
    mutate(
  FitLog = map(data, ~possibnlSLM(OD_600 ~ LogisticEqn(x = E_hours, Intercept, Mu, Pmax),
                            data = .x, 
                            start = list(
                              Intercept = min(.[,OD_600], na.rm = TRUE), 
                              Mu = (log(max(.[,OD_600], na.rm = TRUE)) - log(min(.[OD_600], na.rm = TRUE)))/max(.$E_hours), 
                              Pmax = max(.[,OD_600], na.rm = TRUE)),
                            control = list(maxiter = 1000)
                            ,
                            lower = c((min(.[,OD_600], na.rm = TRUE) * 0.05),
                                      ((log(max(.[, OD_600], na.rm = TRUE)) - log(min(.[,OD_600], na.rm = TRUE)))/max(.$E_hours)) *0.05,
                                      ((max(.[,OD_600], na.rm = TRUE)) * 0.05)),
                            upper = c((min(.[,OD_600], na.rm = TRUE) * 100),
                                      ((log(max(.[,OD_600], na.rm = TRUE)) - log(min(.[,OD_600], na.rm = TRUE)))/max(.$E_hours)) * 100,
                                      ((max(.[,OD_600], na.rm = TRUE)) * 100))
               )
                ),
  PredictLog = map(FitLog, augment),
  TidiedLog = map(FitLog, tidy),
  ParamLog = map(FitLog, glance)
  )





%>%
   mutate(
    FitGompertz = (map(data, ~possibnlSLM(logFitVarminFitVar ~ ModGompertzEqn(x = E_hours, Lag, Gmu, Gmax),
                                    data = .x,
                                    start = list(Lag = LagSeed, Gmu = (log(max(.[, FitVar], na.rm = TRUE)) - log(min(.[ FitVar], na.rm = TRUE)))/max(.$E_hours), Gmax = max(.[, "logFitVarminFitVar"], na.rm = TRUE)),
                                    control = list(maxiter = 1000)
                            ,
                                    lower = c(0, ((log(max(.[, FitVar], na.rm = TRUE)) - log(min(.[ FitVar], na.rm = TRUE)))/max(.$E_hours)) * 0.05, max(.[, "logFitVarminFitVar"], na.rm = TRUE) * 0.05),
                                    upper = c(100, ((log(max(.[, FitVar], na.rm = TRUE)) - log(min(.[ FitVar], na.rm = TRUE)))/max(.$E_hours)) * 100, max(.[, "logFitVarminFitVar"], na.rm = TRUE) * 100)))),
    PredictGompertz = map(FitGompertz, augment),
    TidiedGompertz = map(FitGompertz, tidy),
    ParamGompertz = map (FitGompertz, glance))

TreatLogSpecificPredict <- TreatLogSpecific %>%
  unnest(PredictLog)

TreatLogSpecificCoeff <- TreatLogSpecific %>% 
  unnest(TidiedLog) %>%
  dplyr::select(c(!!sym(RowVar), !!sym(RowVar2), !!sym(NestVar),  !!sym(ColVar), !!sym(ColVar2), !!sym(GroupVar), Wavelength, term:p.value)) %>%
  mutate(Wavelength = as.character(Wavelength))

TreatGompertzSpecificPredict <- TreatLogSpecific %>%
  unnest(PredictGompertz)

TreatGompertzSpecificCoeff <- TreatLogSpecific %>% 
  unnest(TidiedGompertz) %>%
  dplyr::select(c(!!sym(RowVar), !!sym(RowVar2), !!sym(NestVar),  !!sym(ColVar), !!sym(ColVar2), !!sym(GroupVar), Wavelength, term:p.value)) %>%
  mutate(Wavelength = as.character(Wavelength))
```

