---
title: "Plot and Analyze microbial plate growth data"
author: "Maximilian Berthold, Douglas A. Campbell"
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
---

```{r load libraries, echo=FALSE, message = FALSE, warning = FALSE} 
# libraries; Note check actual dependencies
library(tidyverse)
library(purrr)
library(lubridate)
library(googledrive)
library(googlesheets4)
library(stringr)
#library(here)

```

Set variables for project
```{r project data}
Project <- "BIOL3111"
FileIDSingle <- "Ex440"

#deauthorizes access to googlesheet
gs4_deauth()

MetaData <- read_sheet("https://docs.google.com/spreadsheets/d/1vuA18rGk_1h2rdSeEev53VSuqyVLP0JBinbDYp3fCag/edit#gid=0")

DataFolder <- "ProcessData"
DataFit <- "FitData"

#grouping variables for rows & columns, and for nesting for fitting
RowVar <- "FinalPO4_uM"
RowVar2 <- "FinalKNO3_uM"
ColVar <- "FinalGlucose_uM"
ColVar2 <- "FinalCO3_uM"
NestVar <- "FinalFe_uM"
NestVar2 <- "CultureID"
GroupVar <- "Group"

#filtering variables
ChlFl = c("440")
Wavelength_nm <- c("680")
nm_filter <- c("680")

#fitting variable
FitVar <- "RFU"


```


```{r load libraries}
library(tidyverse)
library(lubridate)
library(stringr)
library(broom)
library(knitr)
library(minpack.lm)
library(nlstools)
library(nlsMicrobio)
#library(OneR)

```


```{r read data, warning=FALSE}
ImportGrowth <- readRDS(file = file.path(DataFolder,paste(Project,FileIDSingle, "GrowthLong.Rds", sep  = "_"),fsep = .Platform$file.sep))

#https://stackoverflow.com/questions/64111558/passing-variables-into-the-names-glue-parameter-of-tidyrpivot-wider

GrowthLong <- ImportGrowth %>%
  filter(Wavelength %in% nm_filter) 


GrowthLong <- GrowthLong %>%
  group_by(!!sym(NestVar2)) %>%
  mutate(logFitVarminFitVar = log(!!sym(FitVar)/min(!!sym(FitVar), na.rm = TRUE))) %>% #problem with log leads to NaN, when value is very small
  ungroup()


```


```{r preplot}
GrowthLong %>%
  filter(Wavelength %in% Wavelength_nm) %>%
  ggplot()+
  geom_point(aes(x = E_hours, y = !!sym(FitVar))) +
  facet_grid(cols = vars(!!sym(ColVar))) +
    theme_bw() 
       
FitData <- GrowthLong %>%
  group_by(!!sym(RowVar), !!sym(RowVar2), !!sym(NestVar),  !!sym(ColVar), !!sym(ColVar2), !!sym(GroupVar)) %>%
  mutate(LagSeed = E_hours[which.min(!!sym(FitVar))])

```


Define equations as functions.
x will be taken from 'E_days' when we run the fit.
```{r logistic_eqn}
LogisticEqn <-  function(x, Pmax, Mu, Intercept){(Pmax*Intercept*exp(Mu*x))/(Pmax + (Intercept*(exp(Mu*x)-1)))
}

ModGompertzEqn <- function(x, Gmax,Gmu,Lag){(Gmax*(exp(-exp((Gmu*exp(1))/Gmax*(Lag-x)+1))))}

possibnlSLM = possibly(.f = nlsLM, otherwise = NULL)
```


Fit treatment specific logistic growth trajectories using nest purrr:map & broom::augment using nest specific start, lower & upper settings extracted from each nest on the fly.
This may be necessary if the 'nests' contain diverse data patterns that fail to fit with generic start, lower & upper parameters extracted from the entire dataset


```{r treat growth specific}
#run nesting by passing in values of variables set at top; 
#!!sym(... allows nest to access the characters of the values passed in

TreatNestSpecific <- FitData %>%
  nest(data = -c(!!sym(RowVar), !!sym(RowVar2), !!sym(NestVar),  !!sym(ColVar), !!sym(ColVar2), !!sym(GroupVar), Wavelength, LagSeed))



#extract nest-specific start, lower & upper setting by passing in "."
TreatLogSpecific <- TreatNestSpecific %>% 
    mutate(
  FitLog = map(data, ~possibnlSLM(!!sym(FitVar) ~ LogisticEqn(x = E_hours, Intercept, Mu, Pmax),
                            data = .x, 
                            start = list(
                              Intercept = min(.[,FitVar], na.rm = TRUE), 
                              Mu = (log(max(.[, FitVar], na.rm = TRUE)) - log(min(.[ FitVar], na.rm = TRUE)))/max(.$E_hours), 
                              Pmax = max(.[,FitVar], na.rm = TRUE)),
                            control = list(maxiter = 1000)
                            ,
                            lower = c((min(.[,FitVar], na.rm = TRUE) * 0.05),
                                      ((log(max(.[, FitVar], na.rm = TRUE)) - log(min(.[ FitVar], na.rm = TRUE)))/max(.$E_hours)) *0.05,
                                      ((max(.[,FitVar], na.rm = TRUE)) * 0.05)),
                            upper = c((min(.[,FitVar], na.rm = TRUE) * 100),
                                      ((log(max(.[, FitVar], na.rm = TRUE)) - log(min(.[ FitVar], na.rm = TRUE)))/max(.$E_hours)) * 100,
                                      ((max(.[,FitVar], na.rm = TRUE)) * 100))
               )
                ),
  PredictLog = map(FitLog, augment),
  TidiedLog = map(FitLog, tidy),
  ParamLog = map(FitLog, glance)
  ) %>%
   mutate(
    FitGompertz = (map(data, ~possibnlSLM(logFitVarminFitVar ~ ModGompertzEqn(x = E_hours, Lag, Gmu, Gmax),
                                    data = .x,
                                    start = list(Lag = LagSeed, Gmu = (log(max(.[, FitVar], na.rm = TRUE)) - log(min(.[ FitVar], na.rm = TRUE)))/max(.$E_hours), Gmax = max(.[, "logFitVarminFitVar"], na.rm = TRUE)),
                                    control = list(maxiter = 1000)
                            ,
                                    lower = c(0, ((log(max(.[, FitVar], na.rm = TRUE)) - log(min(.[ FitVar], na.rm = TRUE)))/max(.$E_hours)) * 0.05, max(.[, "logFitVarminFitVar"], na.rm = TRUE) * 0.05),
                                    upper = c(100, ((log(max(.[, FitVar], na.rm = TRUE)) - log(min(.[ FitVar], na.rm = TRUE)))/max(.$E_hours)) * 100, max(.[, "logFitVarminFitVar"], na.rm = TRUE) * 100)))),
    PredictGompertz = map(FitGompertz, augment),
    TidiedGompertz = map(FitGompertz, tidy),
    ParamGompertz = map (FitGompertz, glance))

TreatLogSpecificPredict <- TreatLogSpecific %>%
  unnest(PredictLog)

TreatLogSpecificCoeff <- TreatLogSpecific %>% 
  unnest(TidiedLog) %>%
  dplyr::select(c(!!sym(RowVar), !!sym(RowVar2), !!sym(NestVar),  !!sym(ColVar), !!sym(ColVar2), !!sym(GroupVar), Wavelength, term:p.value)) %>%
  mutate(Wavelength = as.character(Wavelength))

TreatGompertzSpecificPredict <- TreatLogSpecific %>%
  unnest(PredictGompertz)

TreatGompertzSpecificCoeff <- TreatLogSpecific %>% 
  unnest(TidiedGompertz) %>%
  dplyr::select(c(!!sym(RowVar), !!sym(RowVar2), !!sym(NestVar),  !!sym(ColVar), !!sym(ColVar2), !!sym(GroupVar), Wavelength, term:p.value)) %>%
  mutate(Wavelength = as.character(Wavelength))
```



```{r}

Groups = 5

TreatLogSpecificPlot <- TreatLogSpecificPredict %>% 
  filter(Group %in% Groups) %>%
  filter(Wavelength %in% nm_filter,
         FinalKNO3_uM == 40,
         FinalPO4_uM == 80,
         FinalCO3_uM == 80,
         FinalFe_uM == 0.092,
         FinalGlucose_uM == 80
         ) %>%
  ungroup() %>%
  #pivot_longer(cols = c(FinalKNO3_uM:FinalGlucose_uM), names_to = "AddedNutrient", values_to = "NutrientConcentration") %>%
  #filter(NutrientConcentration != 0) %>%
  ggplot() +  
  geom_line(aes(x = E_hours, y = .fitted), size = 0.5) +
  geom_point(aes(x = E_hours, y = !!sym(FitVar)), size = 0.6) +
  # facet_grid(cols = vars(AddedNutrient)
  #            #, rows = vars(NutrientConcentration)
  #            ) +
  theme_bw() +
  labs(y = FitVar,
      title = "Treatment Growth, Specific Log",
      legend.position="top",
      legend.text=element_text(size=20),
      legend.title =element_text(size=20),
      caption = "Log")

TreatLogSpecificPlot

TreatGompertzLagSpecificPlot <- TreatGompertzSpecificPredict %>% 
  filter(Wavelength %in% nm_filter) %>%
  filter(Group %in% Groups) %>%
  # filter(Wavelength %in% nm_filter,
  #        FinalKNO3_uM == 400,
  #        FinalPO4_uM == 8,
  #        FinalCO3_uM == 80,
  #        FinalFe_uM == 0.092,
  #        FinalGlucose_uM == 80
  #        ) %>%
  ungroup() %>%
  pivot_longer(cols = c(FinalKNO3_uM:FinalGlucose_uM), names_to = "AddedNutrient", values_to = "NutrientConcentration") %>%
  filter(NutrientConcentration != 0) %>%
  ggplot() +  
  geom_line(aes(x = E_hours, y = .fitted), size = 0.5) +
  geom_point(aes(x = E_hours, y = logFitVarminFitVar), size = 0.6) +
  facet_grid(cols = vars(AddedNutrient), rows = vars(NutrientConcentration)) +
  theme_bw() +
  labs(y = expression(paste("log(RFU"[`680 nm`]*"/min RFU"[`680 nm`]*")")), x = "Elapsed time (hr)", title = "Modified Gompertz with Lag") +
  theme( legend.position="top",
      legend.text=element_text(size=20),
      legend.title =element_text(size=20),
      axis.text.x=element_text(angle=30, hjust = 1, size=7))

TreatGompertzLagSpecificPlot
```


